---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sfcr

<!-- badges: start -->
<!-- badges: end -->

The goal of `sfcr` is to provide an intuitive and `tidy` way to estimate stock-flow consistent (SFC) models with R. 

With `sfcr`, the models are written entirely with R and use the standard R syntax. Furthermore, their output is a `tibble` that are easily manipulated with the `tidyverse` tools and plotted with `ggplot2`.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("joaomacalos/sfcr")
```
## Example

This is a basic example which shows you how to simulate the steady state of the "SIM" model from @godley2007monetary [ch. 3] and shock this model with an increase in government expenditures.

The first step is to simulate the steady state of the model with the `sfcr_sim()` function:

```{r example}
library(sfcr)
library(ggplot2)

eqs <- list(
  TX_s[t] ~ TX_d[t],
  YD[t] ~ W[t] * N_s[t] - TX_s[t],
  C_d[t] ~ alpha1 * YD[t] + alpha2 * H_h[t-1],
  H_h[t] ~ YD[t] - C_d[t] + H_h[t-1],
  N_s[t] ~ N_d[t],
  N_d[t] ~ Y[t] / W[t],
  C_s[t] ~ C_d[t],
  G_s[t] ~ G_d[t],
  Y[t] ~ C_s[t] + G_s[t],
  TX_d[t] ~ theta * W[t] * N_s[t],
  H_s[t] ~ G_d[t] - TX_d[t] + H_s[t-1]
)

exg <- list("G_d" = 20, "W" = 1)

params <- list("alpha1" = 0.6, "alpha2" = 0.4, "theta" = 0.2)

sim_model <- sfcr_sim(eqs, t = 60, exg, params)
sim_model
```


With the steady state values at hand, we can see what happens if we increase government expenditures ("G_d") from 20 to 30:

```{r}
shock <- list("G_d" = 30)

sim2 <- sfcr_scenario(
  steady_state = sim_model,
  equations = eqs, 
  t = 60, 
  exogenous = exg, 
  parameters = params, 
  shock_exg = shock
  )

sim2
```


The output is conveniently stored as a `tibble`. It means that we can easily plot this model using the `ggplot2` package:

```{r}
sim2 %>%
  ggplot(aes(x = t, y = Y)) +
  geom_line()
```

See the vignettes for developed examples.

### References
