---
title: "Model LP"
output: bookdown::html_fragment2
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Model Liquidity Preference (LP)

This article shows how to implement the model LP from @godley2007monetary [ch. 5].

We start by loading the `sfcr` and the `tidyverse` packages:

```{r setup}
library(sfcr)
library(tidyverse)
```

## Base LP model

The first step is to write down the equations that define the endogenous variables. I will name the equations as it will make it easier to change them later when we move to models "LP2" and "LP3":

```{r}
lp_eqs <-
  list(
    eq1 = Y[t] ~ C[t] + G[t],
    
    eq2 = YD_r[t] ~ Y[t] - TX[t] + r_b[t-1] * B_h[t-1] + BL_h[t-1],
    
    eq3 = TX[t] ~ theta * (Y[t] + r_b[t-1] * B_h[t-1] + BL_h[t-1]),
    
    eq4 = V[t] ~ V[t-1] + (YD_r[t] - C[t]) + CG[t],
    
    eq5 = CG[t] ~ (p_bl[t] - p_bl[t-1]) * BL_h[t-1],
    
    eq6 = C[t] ~ alpha1 * YDE_r[t] + alpha2 * V[t-1],
    
    eq7 = VE[t] ~ V[t-1] + (YDE_r[t] - C[t]) + CG[t],
    
    eq8 = H_h[t] ~ V[t] - B_h[t] - p_bl[t] * BL_h[t],
    
    eq9 = H_d[t] ~ VE[t] - B_d[t] - p_bl[t] * BL_d[t],
    
    eq10 = B_d[t] ~ (VE[t] * lambda20) + VE[t] * (lambda22 * r_b[t] + lambda23 * ERr_bl[t]) + lambda24 * (YDE_r[t]),
    
    eq11 = BL_d[t] ~ VE[t] * (lambda30 + lambda32 * r_b[t] + lambda33 * ERr_bl[t] + lambda34 * (YDE_r[t]/VE[t]))/p_bl[t],
    
    eq12 = B_h[t] ~ B_d[t],
    
    eq13 = BL_h[t] ~ BL_d[t],
    
    eq14 = B_s[t] ~ B_s[t-1] + (G[t] + r_b[t-1] * B_s[t-1] + BL_s[t-1]) - (TX[t] + r_b[t-1] * B_cb[t-1]) - ((BL_s[t] - BL_s[t-1]) * p_bl[t]),
    
    eq15 = H_s[t] ~ H_s[t-1] + (B_cb[t] - B_cb[t-1]),
    
    eq16 = B_cb[t] ~ B_s[t] - B_h[t],
    
    eq17 = BL_s[t] ~ BL_h[t],
    
    eq18 = ERr_bl[t] ~ r_bl[t] + chi * ((pe_bl[t] - p_bl[t])/p_bl[t]),
    
    eq19 = r_bl[t] ~ 1/p_bl[t],
    
    eq20 = pe_bl[t] ~ p_bl[t],
    
    eq21 = CGE[t] ~ chi * (pe_bl[t] - p_bl[t]) * BL_h[t],
    
    eq22 = YDE_r[t] ~ YD_r[t-1]
  )
```

Next we define the parameters and the values for the exogenous variables:

```{r}
lp_exogenous <- list('G' = 20, 'r_b' = 0.03, 'p_bl' = 20)

lp_parameters = list(
  'theta' = 0.1938,
  'alpha1' = 0.8, 
  'alpha2' = 0.2, 
  'lambda20' = 0.44196, 
  'lambda22' = 1.1, 
  'lambda23' = -1,
  'lambda24' = -0.03,
  'lambda30' = .3997,
  'lambda32' = -1,
  'lambda33' = 1.1,
  'lambda34' = -0.03,
  'chi' = 0.1
)
```

And simulate the steady state values:

```{r, paged.print = F}
lp1 <- sfcr_sim(
  equations = lp_eqs, 
  t = 50,
  parameters = lp_parameters, 
  exogenous = lp_exogenous,
  hidden = list("H_s" = "H_h"))

# We check the last values to be sure that we arrived at the steady state with 50 periods:
lp1 %>% tail
```



With the steady state values at hand, we can start adding scenarios to analyze what would happen if we change one of the parameters or exogenous variables.

We start by increasing, at the same time, the short-term and long-term interest rates.


```{r, paged.print = F}
shock_exg <- list("r_b" = 0.04, "p_bl" = 15)

lp_mod_s1 <- sfcr_scenario(
  steady_state = lp_mod,
  equations = equations,
  t = 60,
  exogenous = exogenous,
  parameters = parameters,
  shock_exg = shock_exg
)

lp_mod_s1 %>% tail
```

Before plotting the variables, we need to calculate three different ratios and the total value of bonds from the simulated model:

- Value of bonds $=$ Price of bonds * Bonds held

- Wealth to disposable income ratio

- Bonds to wealth ratio

- Bills to wealth ratio


Here we see how the `tibble` output provides a seamless integration with the `dplyr` tools that were loaded with the `tidyverse` packages:

```{r, paged.print = F}
lp_mod_s1 <-
  lp_mod_s1 %>%
  dplyr::mutate(
    BL = BL_h * p_bl,
    VYDr_r = V / YD_r,
    BhV_r = B_h / V,
    BLV_r = BL / V
  )

# Check the newly created columns:
lp_mod_s1 %>%
  dplyr::select(BL, VYDr_r, BhV_r, BLV_r)
```

We are now ready to plot the output. We use the `ggplot2` package to do so.

The `ggplot2` package works better with data on the long format. Therefore, we must reshape the data before plotting the variables. We will do it with the `tidyr::pivot_longer()` function and save the `tibble` in the long format into a new object.

In the long format, we have one column that store the names of all variables, and only one column for the values. The column `t` records the periods. Note that each period is now represented in 41 rows.


```{r, paged.print = F}
(s1_long <- lp_mod_s1 %>%
  tidyr::pivot_longer(cols = -t, names_to = 'variable'))
```
We are now ready to plot the variables! Let's start with Wealth to disposable income ratio (Figure 5.2 in @godley2007monetary)

The syntax is very simple:

1. We `dplyr::filter` the `variable` column to select only the variables we are interested into;

2. We plot using `ggplot2`. We map the x axis to `t` and the `y` axis to `value`

```{r}
s1_long %>%
  dplyr::filter(variable == 'VYDr_r') %>%
  ggplot(aes(x = t, y = value)) +
  geom_line() +
  labs(x = 'Periods', y = 'Wealth to disposable income ratio')
```

Next, we plot disposable income (YD_r) and consumption (C) (Figure 5.3):

Now, instead of using `==`, we use `%in%` to select two variables. We map `linetype` to `variable` in the `geom_line()` aesthetics to have the multiple lines:

```{r}
s1_long %>%
  filter(variable %in% c("YD_r", "C")) %>%
  ggplot(aes(x = t, y = value)) +
  geom_line(aes(linetype = variable))

```

Finally, we plot the evolution of bonds to wealth ratio and the evolution of bills to wealth ratio:

```{r}
s1_long %>%
  filter(variable %in% c("BLV_r", "BhV_r")) %>%
  ggplot(aes(x = t, y = value)) +
  geom_line(aes(linetype = variable))
```

## LP2

Let's move to the model LP2. This model captures the effect of household liquidity preference on long-term interest rates by changing and addin some of the equations. The reader is invited to check @godley2007monetary for further details on the models and equations

The easiest way to implement make this changes is to save a copy of the `equations` list and modify this new object:

```{r, paged.print = F}
lp2_eqs <- lp_eqs

lp2_eqs$eq20 = pe_bl[t] ~ pe_bl[t-1] - betae * (pe_bl[t-1] - p_bl[t]) + add[t]
lp2_eqs$eq23 = p_bl[t] ~  (1 + zeta1[t] *beta - zeta2[t]*beta) * p_bl[t-1]
lp2_eqs$eq24 = TP[t] ~ (BL_h[t-1] * p_bl[t-1]) / (BL_h[t-1] * p_bl[t-1] + B_h[t-1])
lp2_eqs$eq25 = zeta1[t] ~ if (TP[t] > top) {1} else {0}
lp2_eqs$eq26 = zeta2[t] ~ if (TP[t] < bot) {1} else {0}
#
```


We also need to add the parameters $beta\_e$, $beta$, $top$, and $bot$ to the list of parameters and remove $p\_bl$ from the exogenous variables. 

I also added a bunch of initial values -- taken from Gennaro Zezza's Eviews code -- to get my values as close as possible from the ones presented in the book.


```{r}
lp2_params <- lp_parameters
lp2_params$betae = 0.5
lp2_params$beta = 0.02
lp2_params$top = 0.505
lp2_params$bot = 0.495

lp2_exogenous = list('G' = 20, 'r_b' = 0.03, "add" = 0)

lp2_initial = list(
  "YD_r" = 95.803,
  "V" = 95.803,
  "H_h" = 20.125,
  "BL_d" = 1.89,
  "B_h" = 37.839,
  "BL_h" = 1.892,
  "B_s" = 57.964,
  "H_s" = 20.125,
  "B_cb" = 20.125,
  "BL_s" = 1.89,
  "r_bl" = 0.05,
  "pe_bl" = 20,
  "p_bl" = 20
  )

```

We can simulate the model:

```{r}
lp2 <- sfcr_sim(
  lp2_eqs, 
  t = 60, 
  lp2_exogenous, 
  lp2_params,
  initial = lp2_initial,
  hidden = list("H_h" = "H_s")
  )

lp2 %>% tail()
```

And add a shock to the short interest-rate:

```{r}
shock_exg <- list("r_b" = 0.035)

lp2_s1 <- sfcr_scenario(
  lp2,
  lp2_eqs,
  t = 60,
  exogenous = list('G' = 20, 'r_b' = 0.03, "add" = 0),
  lp2_params,
  shock_exg = list("r_b" = 0.035),
  )
```

As before, we create a new `tibble` with the model in long format:

```{r}
lp2_s1_long <- lp2_s1 %>%
  pivot_longer(cols = -t, names_to = "variable")
```

And plot the variables. We start by plotting the evolution of the share of bonds in government debt held by households (TP), as in Figure 5.6:

```{r}
lp2_s1_long %>%
  filter(variable == 'TP',
         t < 51) %>%
  ggplot(aes(x = t, y = value)) +
  geom_line() +
  geom_hline(yintercept = 0.505) +
  geom_hline(yintercept = 0.495)
```
We also can visualize the evolution of interest rates (Figure 5.5):

```{r}
lp2_s1_long %>%
  filter(variable %in% c("r_b", "r_bl"),
         t < 51) %>%
  ggplot(aes(x = t, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = "free_y")
```
What would happen if households anticipated a fall in the price of bonds? To see it, we need to create a new scenario adding a one-time shock to the $add$ component of equation 20.

We do it by filling the `one_time_shock` argument of `sfcr_scenario()`. This argument must be a named list, with the name of the exogenous variable being shocked as the element' name, and a vector containing the one-time shock and the period that it should take place.

In the example below, we add the line `one_time_shock = list("add" = c(-2.8, 6))` to the `sfcr_scenario()` call to model this shock:

```{r}
lp2_s2 <- sfcr_scenario(
  lp2,
  lp2_eqs,
  t = 60,
  exogenous = list('G' = 20, 'r_b' = 0.03, "add" = 0),
  lp2_params,
  one_time_shock = list("add" = c(-2.5, 6))
  )
```

We create a "long" `tibble`:

```{r}
lp2_s2_long <- lp2_s2 %>%
  pivot_longer(cols = -t, names_to = "variable")
```

In order to plot the variables.

We start by plotting the evolution of the price of bonds ($p\_bl$) and the expected price of bonds ($pe\_bl$), as in Figure 5.8:

```{r}
lp2_s2_long %>%
  filter(variable %in% c("p_bl", "pe_bl")) %>%
  ggplot(aes(x = t, y = value)) +
  geom_line(aes(linetype = variable))
```
This shock to the expected price of bonds cause the long-term interest rate to rise, as a consequence of the target proportion of bonds in the total government debt aimed by the government:

```{r}
lp2_s2_long %>%
  filter(variable == "TP") %>%
  ggplot(aes(x = t, y = value)) +
  geom_line(linetype = 2) +
  geom_hline(yintercept = 0.505) +
  geom_hline(yintercept = 0.495) +
  annotate("text", x = 15, y = 0.490, label = "Share of bonds in \ngovernemnt debt \nheld by households")
```
```{r}
lp2_s2_long %>%
  filter(variable == "r_bl") %>%
  ggplot(aes(x = t, y = value)) +
  geom_line(linetype = 2) +
  annotate("text", x = 25, y = 0.05195,
           label = "Long-term interest rate")
```
## LP3: Making government expenditures endogenous

To simulate the model LP3, we need to add four new equations to the model LP2. Here, it is easier to find the steady state of the model keeping $G$ as a exogenous variable:

```{r}
lp3_eqs <- lp2_eqs

lp3_eqs$eq27 <- PSBR[t] ~ (G[t] + r_b[t] * B_s[t-1] + BL_s[t-1]) - (TX[t] + r_b[t] * B_cb[t-1])
lp3_eqs$eq28 <- zeta3 ~ if ( (PSBR[t-1]/Y[t-1]) > 0.03 ) {1} else {0}
lp3_eqs$eq29 <- zeta4 ~ if ( (PSBR[t-1]/Y[t-1]) < -0.03 ) {1} else {0}
```

```{r}
lp3_params <- lp2_params
lp3_params$beta_g <- 1

lp3_initial <- lp2_initial

lp3 <- sfcr_sim(
  lp3_eqs,
  t = 80,
  exogenous = list('r_b' = 0.03, "add" = 0, "add2" = 0, "G" = 20),
  parameters = lp3_params,
  initial = lp3_initial,
  hidden = c("H_s" = "H_h"))

lp3 %>% select(t, G, PSBR, everything())
```
What would happen in this model if there was a sharp reduction in the propensity to consume? That's the objective of this exercise with model LP3.

First, we add the equation that endogenize $G$ to the `lp3_eqs` list, and after we calculate this scenario:

```{r}
lp3_eqs$eq30 <- G[t] ~ G[t-1] - (zeta3 + zeta4) * beta_g * PSBR[t-1] + add2[t]

lp3_s1 <- sfcr_scenario(
  lp3,
  lp3_eqs,
  t = 60,
  exogenous = list("r_b" = 0.03, "add" = 0, "add2" = 0),
  parameters = lp3_params,
  shock_param = list("alpha1" = 0.7)
) %>% mutate(
  PSBR_Y = PSBR / Y
)

lp3_s1_long <- lp3_s1 %>% pivot_longer(cols = -t, names_to = "variable")
```

Here we can see that the fall in output is permanent:

```{r}
lp3_s1_long %>%
  filter(variable == "Y") %>%
  ggplot(aes(x = t, y = value)) +
  geom_line()
```
This result contrasts with the long-run effect of a reduction in the propensity to consume arrived with the LP model:

```{r}
lp1_s2 <- sfcr_scenario(
  lp1,
  lp_eqs,
  t = 60,
  lp_exogenous,
  lp_parameters,
  shock_param = list("alpha1" = 0.7)
)

lp1_s2_long <- lp1_s2 %>% pivot_longer(cols = -t, names_to = "variable")
```

```{r}
lp1_s2_long %>%
  filter(variable == "Y") %>%
  ggplot(aes(x = t, y = value)) +
  geom_line()
```
In the LP model, the steady-state output level is higher than initially because the consumption out of the wealth ends up more than compensanting the initial fall in output.

In model LP3, however, the initial fall in output increases the government deficit, and this causes the government to reduce its expenditures:

```{r}
lp3_s1_long %>%
  filter(variable %in% c("G", "PSBR_Y"),
         t < 51) %>%
  ggplot(aes(x = t, y = value)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_y")
```

## Sankey's diagram

A Sankey's diagram is used in the context of SFC models to visualize the transactions flow matrix. Hence, its important to check this matrix first before starting the diagram:

```{r, paged.print = F}
(lp_tfm <- tribble(
  ~variable, ~hh, ~prod, ~gvt, ~cb_c, ~cb_k,
  "Consumption", "-C", "C", NA, NA, NA,
  "Gvt Expenditures", NA, "G", "-G", NA, NA,
  "GDP", "Y", "-Y", NA, NA, NA,
  "Int. on bills", "r_b-1 * B_h-1", NA, "- r_b-1 * B-1", "r_b-1 * B_cb-1", NA,
  "Int. on bonds", "BL-1", NA, "-BL-1", NA, NA,
  "CB profits", NA, NA, "r_b-1 * B_cb-1", "-r_b-1 * B_cb-1", NA,
  "Taxes", "-TX", NA, "TX", NA, NA,
  "Change money", "-delta H", NA, NA, NA, "delta H",
  "Change bills", "-delta B_h", NA, "delta B", NA, "-delta B_cb",
  "Change bonds", "-delta BL * p_bl", NA, "delta BL * p_bl", NA, NA,
  "Capital Gains", "-delta p_bl * BL-1", NA, "delta p_bl * BL-1", NA, NA,
) %>%
  mutate(row = row_number())) #%>%
  #knitr::kable() %>%
  #kableExtra::kable_styling(full_width = F)
```
Now, we have need the entries in each row as a variable. Therefore, we need to make some new columns:

```{r, paged.print = F}
(lp_sankey <- lp1 %>%
  mutate(
    int_bills_h = lag(r_b) * lag(B_h),
    int_bills_g = lag(r_b) * lag(B_s),
    int_bills_cb = lag(r_b) * lag(B_cb),
    int_bonds = lag(BL_h),
    cb_profits = lag(r_b) * lag(B_cb),
    delta_h = H_h - lag(H_h),
    delta_b_h = B_h - lag(B_h),
    delta_b_g = B_s - lag(B_s),
    delta_b_cb = B_cb - lag(B_cb),
    delta_BL = BL_h - lag(BL_h),
    delta_BL_P = delta_BL * p_bl) %>%
  select(t, C, G, Y, int_bills_h, int_bills_g, int_bills_cb, int_bonds, cb_profits,
         TX, delta_h, delta_b_h, delta_b_g, delta_b_cb, delta_BL_P) %>%
  slice_tail(n = 1))
#
```

Next, we convert the model to the long format, and associate each entry with its row in the TFM:

```{r}
(lp_sankey2 <- lp_sankey %>% 
  pivot_longer(cols = -t) %>%
  mutate(row = c(1, 2, 3, 4, 4, 4, 5, 6, 7, 8, 9, 9, 9, 10)) %>%
   select(-t))
#
```


Next, we calculate two `tibble`s for each variable on the TFM. One with the positive entries and one with the negative entries, and we create a column for the sector that each of this entry corresponds. This column has the following identification:

```{r}
sector_map <-
  tribble(
    ~sector_name, ~sector,
    'hh', 1,
    'pr', 2,
    'gvt', 3,
    'cb_c', 4,
    'cb_k', 5
  )
```


```{r, paged.print = F}
# Positive
lp_sankey_plus <- lp_sankey2 %>%
  filter(!str_detect(name, "bills_g"),
         !str_detect(name, "b_h|b_cb")) %>%
  mutate(sector = c(2, 2, 1, 1, 4, 1, 3, 3, 5, 3, 3)) %>%
  mutate(name = if_else(str_detect(name, "bills"), "int_bills", name),
         name = if_else(str_detect(name, 'delta_b'), 'delta_b', name)) %>%
  left_join(sector_map)

(lp_sankey_minus <- lp_sankey2 %>%
  filter(!str_detect(name, "bills_h|bills_cb"),
         !str_detect(name, "b_g")) %>%
  mutate(sector = c(1, 3, 2, 3, 3, 4, 1, 1, 1, 5, 1)) %>%
  mutate(name = if_else(str_detect(name, "bills"), "int_bills", name),
         name = if_else(str_detect(name, 'delta_b'), 'delta_b', name)) %>%
  left_join(sector_map))
```

Create nodes `tibble`:

```{r}
(lp_nodes <- tibble(
  name = c(unique(lp_sankey_plus$name),
           unique(lp_sankey_plus$sector_name),
           str_c(unique(lp_sankey_plus$sector_name), 1)),
  node = 0:19
))
```

Create links `tibble`:

```{r}
links_lp_minus <- lp_sankey_minus %>%
  left_join(lp_nodes, by = c('name')) %>%
  left_join(lp_nodes, by = c("sector_name" = "name")) %>%
  rename(source = node.y, target = node.x)

links_lp_plus <- lp_sankey_plus %>%
  mutate(sector_name = str_c(sector_name, "1")) %>%
  left_join(lp_nodes, by = c('name')) %>%
  left_join(lp_nodes, by = c("sector_name" = "name")) %>%
  rename(source = node.x, target = node.y)
```

Combine the two datasets, select the relevant columns, and rename:

```{r}
links_lp <- bind_rows(links_lp_minus, links_lp_plus) %>%
  select(source, target, value)
```

Plot the Sankey's diagram:

```{r}
lp_nodes1 <- lp_nodes %>%
  mutate(name = case_when(
    name == "C" ~ "Consumption",
    name == "G" ~ "Gvt. expenditures",
    name == "Y" ~ "Output",
    name == "int_bills" ~ "Interest payments on bills",
    name == "int_bonds" ~ "Interest payments on bonds",
    name == "cb_profits" ~ "Central Bank profits",
    name == "TX" ~ "Taxes",
    name == "delta_h" ~ "\u0394 monetary balances",
    name == "delta_b" ~ "\u0394 bills",
    name == "delta_BL" ~ "\u0394 bills",
    str_detect(name, 'pr') ~ "Firms",
    str_detect(name, 'gvt') ~ "Government",
    str_detect(name, 'cb_c') ~ "Central Bank current",
    str_detect(name, 'cb_k') ~ "Central Bank capital",
    str_detect(name, "hh") ~ "Households"
  ))

networkD3::sankeyNetwork(Links = as.data.frame(links_lp), Nodes = as.data.frame(lp_nodes1), 
                          Source = 'source', 
                          Target = 'target', 
                          Value = 'value', 
                          NodeID = 'name',
                          units = 'dollars',
                          fontSize = 14)
```

