---
title: "Chapter 3: Model SIM"
output: 
  rmarkdown::html_document:
    toc: true
references:
- id: godley2007monetary
  title: >
       Monetary Economics: An Integrated Approach To Credit, Money, Income, Production and Wealth
  author:
  - family: Godley
    given: Wynne
  - family: Lavoie
    given: Marc
  publisher: Palgrave Macmillan
  type: book
  issued:
    year: 2007
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  paged.print = F
)
```

In this article we show how the `sfcr` package can be used to replicate the model SIM from @godley2007monetary [ch. 3].

We start by loading the `sfcr` package and the `tidyverse` packages, that are going to be used to manipulate the data and generate the plots:

```{r setup}
library(sfcr)
library(tidyverse)
```


# SIM Model

## Equations

The first step is to write down the equations, exogenous variables, and parameters:

```{r}
sim_eqs <- sfcr_set(
  TXs ~ TXd,
  YD ~ W * Ns - TXs,
  Cd ~ alpha1 * YD + alpha2 * Hh[-1],
  Hh ~ YD - Cd + Hh[-1],
  Ns ~ Nd,
  Nd ~ Y / W,
  Cs ~ Cd,
  Gs ~ Gd,
  Y ~ Cs + Gs,
  TXd ~ theta * W * Ns,
  Hs ~ Gd - TXd + Hs[-1]
)

sim_ext <- sfcr_set(
  Gd ~ 20,
  W ~ 1,
  alpha1 ~ 0.6,
  alpha2 ~ 0.4,
  theta ~ 0.2
)
```


## Baseline

With these objects defined, we can simulate the SIM model. We will make it run for 100 periods. We will also define the hidden equation beforehand, ensuring that the model is written correctly:

```{r}
system.time(sim <- sfcr_baseline(
  equations = sim_eqs, 
  external = sim_ext, 
  periods = 100, 
  hidden = c("Hh" = "Hs"),
  method = "Broyden")
  )

```

We can check that the model generates the same long-run results as the model presented in page 69 @godley2007monetary..

```{r}
sim %>%
  filter(period %in% c(1, 2, 3, 50)) %>%
  select(period, Gs, Y, TXd, YD, Hs) %>%
  t() %>%
  round(digits = 0)
  
```



### A permanent increase in government expenditures

What happens if there is a permanent increase in government expenditures? To check it, we need to use the `sfcr_shock()` and `sfcr_scenario()` functions:

```{r}
shock1 <- sfcr_shock(
  variables = list(
    Gd ~ 25
  ),
  start = 5,
  end = 50
)

sim2 <- sfcr_scenario(sim, shock1, 50)
```

To visualize the consequences of the shock with the `ggplot2` package, we need to first transform the model into long format. We can do it with `pivot_longer()` function from `dplyr`. 

The syntax might look complicated but it is actually simple. With the `pivot_longer()` function we want to transform all columns from the `sim2` object into the long format, indexed by the `period` column. That's why in cols we "remove" the period column.

To plot, we filter the `sim2_long` object to keep only $Y$. As we can see, an increase in government leads to an increase in production in the long run.

```{r}
sim2_long <- sim2 %>%
  pivot_longer(cols = -period)

sim2_long %>%
  filter(name == "Y") %>%
  ggplot(aes(x = period, y = value)) +
  geom_line()
```

To plot multiple variables in the same plot, we need to filter not for a specific variable, but for a group of variables. We do it by using the R operator `%in%` instead of `==`. 

Let's see how to plot $YD$, $Cd$, and $Hh$:

```{r}
sim2_long %>%
  filter(name %in% c("YD", "Cd", "Hh")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line(aes(linetype = name, color = name))
```


# SIMEX model

The SIMEX model expands model SIM by introducing expectations into the model. Technically, it modifies one of the equations and add two new equations.

We use the `sfcr_set()` function as well to modify a set of equations. To modify an existing equation, we must write its new version and exclude the old one from the set with the `exclude` argument. To find the correct id to exclude, we can use the `sfcr_set_index()` function and then `filter` the `lhs` of the original set:

```{r}
sfcr_set_index(sim_eqs) %>%
  filter(lhs == "Cd")

simex_eqs <- sfcr_set(
  sim_eqs,
  Cd ~ alpha1 * YDE + alpha2 * Hh[-1],
  Hd ~ Hd[-1] + YDE - Cd,
  YDE ~ YD[-1],
  exclude = 3
)
```

We use these new equations to simulate the SIMEX model:

```{r}
simex <- sfcr_baseline(simex_eqs, sim_ext, 50, hidden = c("Hh" = "Hs"))
```

And immediately add a shock to propensity to consume in this model:

```{r}
shock2 <- sfcr_shock(
  variables = sfcr_set(alpha1 ~ 0.7),
  start = 5,
  end = 50
)

simex2 <- sfcr_scenario(simex, shock2, 50)
```

We then wrap up this article by plotting the consequences of an increase in the propensity to consume on consumption, disposable income, and wealth:

```{r}
simex2_long <- simex2 %>%
  pivot_longer(cols = -period)

simex2_long %>%
  filter(name %in% c("Cd", "YD", "Hh")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line(aes(linetype = name))
```

An attentive reader would notice that the Figure above is not the same as Figure 3.8 of @godley2007monetary. That's because Figure 3.8 shocks the SIM model and not the SIMEX model. I'll leave as an exercise for the reader to generate Figure 3.8.

# Extra functionalities of the `sfcr` package

## Matrices of model SIM

The `sfcr` package also provides functions to write and validate the balance-sheet and transactions-flow matrices of any SFC model. It further ensures that the model is stock-flow consistent. Here, the main functions are `sfcr_matrix()`, to write the matrices, and `sfcr_validate()`, to check whether the matrix is consistent with the simulated model.

The balance-sheet matrix of model SIM is too simple, however, and not worth presenting here as it would require us to create two extra variables that are missing variables: the wealth of households, and public debt. These entries are important because all columns in the balance-sheet matrix must sum to zero.

### Transactions-flow matrix

```{r}
tfm_sim <- sfcr_matrix(
  columns = c("Households", "Firms", "Government"),
  codes = c("h", "f", "g"),
  c("Consumption", h = "-Cd", f = "+Cs"),
  c("Govt. Exp.", f = "+Gs", g = "-Gd"),
  c("Factor Income", h = "W * Ns", f = "-W * Ns"),
  c("Taxes", h = "-TXs", g = "+TXd"),
  c("Ch. Money", h = "-d(Hh)", g = "d(Hs)")
)
```


```{r}
sfcr_validate(tfm_sim, sim, which = "tfm")
```

### Sankey's diagram

Another benefit of having a transactions-flow matrix that is validated in the model is that we can confidently construct a Sankey's representation of this matrix with the `sfcr_sankey()` function:

```{r}
sfcr_sankey(tfm_sim, sim)
```

## DAG: The structure of the model

The `sfcr` package also provides a built-in function to visualize the DAG representation of the equations in the model:

```{r}
sfcr_dag_cycles_plot(sim_eqs, size = 10)
```


## Blocks of independent equations

If the user is interested, she or he can check how many iterations was necessary per block to converge to a result by calling the `sfcr_get_matrix()` function on the `sim` object. 

```{r}
sim_matrix <- sfcr_get_matrix(sim)

round(head(sim_matrix), digits = 2)
```

To check which variables are included in each block, the user should use the `sfcr_get_blocks()` function:

```{r}
sfcr_get_blocks(sim)
```

It is also possible to visualize the block structure of the model with the `sfcr_dag_blocks_plot()` function:

```{r}
sfcr_dag_blocks_plot(sim_eqs)
```

All of the solvers included in the `sfcr` package first organize the model in blocks, using the two depth-first searches provided by the `igraph` package. The simultaneous equations solvers ("Gauss", "Broyden", or "Newton") are only called to solve for the values of the cyclical variables in the model.

# References



