---
title: "Chapter 7: Model BMW"
output: 
  rmarkdown::html_document:
    toc: true
references:
- id: godley2007monetary
  title: >
       Monetary Economics: An Integrated Approach To Credit, Money, Income, Production and Wealth
  author:
  - family: Godley
    given: Wynne
  - family: Lavoie
    given: Marc
  publisher: Palgrave Macmillan
  type: book
  issued:
    year: 2007
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Load the required packages for this analysis:

```{r setup}
library(sfcr)
library(tidyverse)
```

# Bank-Money World model

## Equations

1. Write down the equations of the model:

```{r}
bmw_eqs <- sfcr_set(
  # Basic behavioral equations
  Cs ~ Cd,
  Is ~ Id,
  Ns ~ Nd,
  Ls ~ Ls[-1] + Ld - Ld[-1],
  
  # Transactions of the firms
  Y ~ Cs + Is,
  WBd ~ Y - rl[-1] * Ld[-1] - AF,
  AF ~ delta * K[-1],
  Ld ~ Ld[-1] + Id - AF,
  
  # Transactions of households
  YD ~ WBs + rm[-1] * Mh[-1],
  Mh ~ Mh[-1] + YD - Cd,
  
  # Transactions of the banks
  Ms ~ Ms[-1] + Ls - Ls[-1],
  rm ~ rl,
  
  # The wage bill
  WBs ~ W * Ns,
  Nd ~ Y / pr,
  W ~ WBd / Nd,
  
  # Household behavior
  Cd ~ alpha0 + alpha1 * YD + alpha2 * Mh[-1],
  
  # The investment behavior
  K ~ K[-1] + Id - DA,
  DA ~ delta * K[-1],
  KT ~ kappa * Y[-1],
  Id ~ gamma * (KT - K[-1]) + DA
)

bmw_external <- sfcr_set(
  rl ~ 0.025,
  alpha0 ~ 20,
  alpha1 ~ 0.75,
  alpha2 ~ 0.10,
  delta ~ 0.10,
  gamma ~ 0.15,
  kappa ~ 1,
  pr ~ 1
)
```

## Baseline

This model is tricky. The "Broyden" does not converge to a sensible answer if we don't set the initial values of the following variables to reasonably large values (as the ones below). Interestingly, the other two solvers ("Gauss" and "Newton") succeed in arriving to a sensible steady state without setting the initial values:

```{r}
bmw_initial <- sfcr_set(
    Ns ~ 128, # Needs to be set
    Ls ~ 12, # Needs to be set
    Y ~ 128, # Needs to be set
    Ld ~ 12, # Needs to be set
    Mh ~ 12, # Needs to be set
    Ms ~ 12, # Needs to be set
    Nd ~ 128, # Needs to be set
    )

bmw <- sfcr_baseline(bmw_eqs, bmw_external, periods = 100, initial = bmw_initial, method = "Broyden", hidden = c("Ms" = "Mh"))
```


```{r}
bmw <- sfcr_baseline(bmw_eqs, bmw_external, periods = 100, method = "Newton", hidden = c("Ms" = "Mh"))
```


## DAG

Let's check the internal structure of this model with the `sfcr_dag_blocks_plot()` function:

```{r}
sfcr_dag_blocks_plot(bmw_eqs)
```

First, let's check that the model arrived to a steady state income level:


```{r}
bmw %>%
  pivot_longer(cols = -period) %>%
  filter(name == "Y") %>%
  ggplot(aes(x = period, y = value)) +
  geom_line() +
  labs(title = "Steady state GDP")
```

## Matrices of model BMW

### Balance-sheet

```{r}
bs_bmw <- sfcr_matrix(
  columns = c("Households", "Firms", "Banks", "Sum"),
  codes = c("h", "f", "b", "s"),
  c("Money", h = "+Mh", b = "-Ms"),
  c("Loans", f = "-Ld", b = "+Ls"),
  c("Fixed capital", f = "+K", s = "+K"),
  c("Balance", h = "-Mh", s = "-Mh")
)
```

Validate:
```{r}
sfcr_validate(bs_bmw, bmw, "bs")
```

### Transactions-flow

```{r}
tfm_bmw <- sfcr_matrix(
  columns = c("Households", "Firms_current", "Firms_capital", "Banks_current", "Banks_capital"),
  codes = c("h", "fc", "fk", "bc", "bk"),
  c("Consumption", h ="-Cs", fc = "+Cd"),
  c("Investment", fc = "+Is", fk = "-Id"),
  c("Wages", h = "+WBs", fc = "-WBd"),
  c("Depreciation", fc = "-AF", fk = "+AF"),
  c("Interest loans", fc = "-rl[-1] * Ld[-1]", bc = "+rl[-1] * Ls[-1]"),
  c("Interest on deposits", h = "+rm[-1] * Mh[-1]", bc = "-rm[-1] * Ms[-1]"),
  c("Ch. loans", fk = "+d(Ld)", bk = "-d(Ls)"),
  c("Ch. deposits", h = "-d(Mh)", bk = "+d(Ms)")
)
```


```{r}
sfcr_validate(tfm_bmw, bmw, "tfm")
```
Now we add some shocks to this model to see how it behaves in different scenarios.

## Scenario 1: An increase in autonomous consumption expenditures

First, let's write a helper function to convert the models to the long format, filter the variables we want to plot, and make a line plot:

```{r}
do_plot <- function(model, variables, title = NULL) {
  model %>%
    mutate(YK = Y / lag(K)) %>%
    pivot_longer(cols = -period) %>%
    filter(name %in% variables) %>%
    ggplot(aes(x = period, y = value)) +
    geom_line(aes(linetype = name)) +
    labs(title = title)
}
```

We then write the shock we want to add and simulate the new scenario:

```{r}
shock1 <- sfcr_shock(
  variables = sfcr_set(
    alpha0 ~ 30
  ),
  start = 5,
  end = 60
)

bmw1 <- sfcr_scenario(bmw,
                      scenario = shock1,
                      periods = 60)
```

How the consumption of households and their disposable income evolves?

```{r}
bmw1 %>%
  do_plot(variables = c("Cd", "YD"), 
          title = "BMW1: Evolution of consumption and disposable income")
```

And the investment of firms?

```{r}
bmw1 %>%
  do_plot(variables = c("Id", "AF"),
          title = "BMW1: Investment of firms")
```

## Increase in the propensity to save

```{r}
shock2 <- sfcr_shock(
  variables = sfcr_set(
    alpha1 ~ 0.7
  ),
  start = 5,
  end = 60
)

bmw2 <- sfcr_scenario(bmw,
                      scenario = shock2,
                      periods = 60)
```


```{r}
bmw2 %>%
  do_plot(variables = c("Cd", "YD"), title = "BMW2: Decrease in the propensity to save")
```
 As can be seen, a decrease in the propensity to save in this model leads to a permanent decrease in consumption and on disposable income.
 
```{r}
bmw2 %>%
  do_plot(variables = c("YK"), title = "BMW2: Decrease in the propensity to save")
```
As a note, it is possible to add `ggplot2` arguments to our helper function:

```{r}
bmw2 %>%
  do_plot(variables = c("W"), title = "BMW2: Decrease in the propensity to save") +
  labs(subtitle = "Evolution of the real wage")
```

# Model BMWK

```{r}
sfcr_set_index(bmw_eqs) %>%
  dplyr::filter(lhs == "Cd")

bmwk_eqs <- sfcr_set(
  bmw_eqs,
  Cd ~ alpha0 + alpha1w * WBs + alpha1r * rm[-1] * Mh[-1] + alpha2 * Mh,
  exclude = 16)

sfcr_set_index(bmw_external) %>%
  dplyr::filter(lhs == "alpha1")
  
bmwk_ext <- sfcr_set(
  bmw_external,
  alpha1w ~ 0.8,
  alpha1r ~ 0.15,
  exclude = 3
  
)
```


```{r}
bmwk <- sfcr_baseline(equations = bmwk_eqs,
                 external = bmwk_ext,
                 periods = 100,
                 hidden = c("Ms" = "Mh"),
                 tol = 1e-20,
                 method = "Newton"
                 )
```

```{r}
bmwk %>%
  do_plot(variables = 'Y', "BMWK: Steady state")
```

Let's see the impact of an increase on interest rates in this model:

```{r}
shock3 <- sfcr_shock(
  variables = sfcr_set(rl ~ 0.035),
  start = 5,
  end = 100
)

bmwk1 <- sfcr_scenario(bmwk, shock3, 100, tol = 1e-10)

bmwk1 %>%
  do_plot("Y", "BMWK1: Increase in the interest rate (rm)") +
  labs(subtitle = "Evolution of Gross Domestic Income (Y)")
```

This model has a very interesting characteristic. If we don't strive for numerical precision with a very small `tol` argument at the `bmw` model, the tiny differences in each period lead the model to explosion:

```{r}
shock3 <- sfcr_shock(
  variables = sfcr_set(rl ~ 0.035),
  start = 5,
  end = 200
)

bmwk1_explode <- sfcr_scenario(bmwk, shock3, 200, tol = 1e-5, method = "Newton")

bmwk1_explode %>%
  do_plot("Y", "BMWK1_EXP: Increase in the interest rate (rm)") +
  labs(subtitle = "Explosion of Gross Domestic Income (Y)")
```

# Stability tests

Let's try to visualize the sensitivity of the model BMW to different parameter values, as discussed in @godley2007monetary [section 7.5.2]

We'll start by expanding the values of the `kappa` parameter in the external set:

```{r}
bmw_xext <- sfcr_expand(bmw_external, kappa, seq(1, 2.6, 0.2))
xbmw <- sfcr_multis(bmw_xext, bmw_eqs, periods = 60, method = "Newton")
```

We can see that the model start to become very unstable from $kappa > 1.6$:

```{r, fig.width=9}
getPalette <- grDevices::colorRampPalette(RColorBrewer::brewer.pal(8, "Dark2"))

bind_rows(xbmw[c(1, 2, 3, 4, 5)]) %>%
  mutate(simulation = as_factor(simulation)) %>%
  pivot_longer(cols = -c(period, simulation)) %>%
  filter(name %in% c("Y", "YD", "C")) %>%
  ggplot(aes(x = period, y = value, color = simulation)) +
  geom_line(size = 1, alpha = 1) +
  theme(legend.position = "bottom") +
  scale_color_manual("kappa",
     values = getPalette(20),
     labels = as.character(seq(1, 2.6, 0.2))) +
  facet_wrap(~ name, scales = 'free_y')
```

# References














