---
title: "Portfolio Choice Model"
references:
- id: godley2007monetary
  title: >
       Monetary Economics: An Integrated Approach To Credit, Money, Income, Production and Wealth
  author:
  - family: Godley
    given: Wynne
  - family: Lavoie
    given: Marc
  publisher: Palgrave Macmillan
  type: book
  issued:
    year: 2007
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  paged.print = F
)
```

```{r setup}
library(tidyverse)
library(sfcr)
```

# The matrices of model PC

We start the analysis by writing down the balance-sheet and transactions-flow matrices of the model:

```{r}
bs_pc <- sfcr_matrix(
  columns = c("Households", "Firms", "Government", "Central bank", "sum"),
  codes = c("h", "f", "g", "cb", "s"),
  r1 = c("Money", h = "+Hh", cb = "-Hs"),
  r2 = c("Bills", h = "+Bh", g = "-Bs", cb = "+Bcb"),
  r3 = c("Balance", h = "-V", g = "+V")
)

sfcr_matrix_display(bs_pc, "bs")
```

```{r}
tfm_pc <- sfcr_matrix(
  columns = c("Households", "Firms", "Government", "CB current", "CB capital"),
  codes = c("h", "f", "g", "cbc", "cbk"),
  c("Consumption", h = "-C", f = "+C"),
  c("Govt. Expenditures", f = "+G", g = "-G"),
  c("Income", h = "+Y", f = "-Y"),
  c("Int. payments", h = "+r[-1] * Bh[-1]", g = "-r[-1] * Bs[-1]", cbc = "+r[-1] * Bcb[-1]"),
  c("CB profits", g = "+r[-1] * Bcb[-1]", cbc = "-r[-1] * Bcb[-1]"),
  c("Taxes", h = "-TX", g = "+TX"),
  c("Ch. Money", h = "-(Hh - Hh[-1])", cbk = "+(Hs - Hs[-1])"),
  c("Ch. Bills", h = "-(Bh - Bh[-1])", g = "+(Bs - Bs[-1])", cbk = "-(Bcb - Bcb[-1])")
)

sfcr_matrix_display(tfm_pc)
```

## The model

To simulate the model, we first write down the equations and the external values:

```{r}
pc_eqs <- sfcr_set(
  Y ~ C + G,
  YD ~ Y - TX + r[-1] * Bh[-1],
  TX ~ theta * (Y + r[-1] * Bh[-1]),
  V ~ V[-1] + (YD - C),
  C ~ alpha1 * YD + alpha2 * V[-1],
  Hh ~ V - Bh,
  Hh1 ~ V * ((1 - lambda0) - lambda1 * r + lambda2 * ( YD/V )), # EQ 4.6A
  Bh ~ V * (lambda0 + lambda1 * r - lambda2 * ( YD/V )),
  Bs ~ Bs[-1] + (G + r[-1] * Bs[-1]) - (TX + r[-1] * Bcb[-1]),
  Hs ~ Hs[-1] + Bcb - Bcb[-1],
  Bcb ~ Bs - Bh
)

pc_ext <- sfcr_set(
  # Exogenous
  
  r ~ 0.025,
  G ~ 20,
  
  # Parameters
  
  alpha1 ~ 0.6,
  alpha2 ~ 0.4,
  theta ~ 0.2,
  lambda0 ~ 0.635,
  lambda1 ~ 0.05,
  lambda2 ~ 0.01
)

```


We then simulate the model to get the steady state values:

```{r}
pc <- sfcr_baseline(
  equations = pc_eqs, 
  external = pc_ext, 
  periods = 70, 
  hidden = c("Hh" = "Hs")
  )
```

The fact that the hidden equation is fulfilled already indicates that the accounting part of the model is water tight. However, the `sfcr` package also allow the user to validate the model using the balance-sheet and transactions-flow matrices.

```{r}
sfcr_validate(bs_pc, pc, "bs")
```
```{r}
sfcr_validate(tfm_pc, pc, "tfm")
```

Given the portfolio choice equations, the model is well specified if equation 4.6A is equal to 4.6. We can use the base `all.equal()` function to check it:

```{r}
all.equal(pc$Hh, pc$Hh1)
```

We can safely conclude that the baseline model is well specified. The `sfcr` package also allows the user to plot the DAG representation of the model and a Sankey's diagram of the transactions-flow matrix. Let's check it out:

## DAG

Let's  use the `sfcr_dag_cycles_plot()` function to see the structure of the model:

```{r}
sfcr_dag_cycles_plot(pc_eqs, size = 10)
```

As can be seen, there's a cycle in this model. The `sfcr` package relies on the `igraph` package to identify the best ordering of the equations to apply the Gauss-Seidel algorithm, reducing the the its computational burden. The blocks of independent equations and the iterations needed at each block to simulate the model can be retrieved by calling the attribute `matrix` of the model.

Alternatively, it is also possible to use the `sfcr_dag_blocks_plot()` function to visualize this structure:

```{r}
sfcr_dag_blocks_plot(pc_eqs, size = 10)
```

## Sankey's diagram

Another functionality offered by the `sfcr` package is that it can generate a Sankey's diagram representation of the transactions-flow matrix.

Here, it is crucial to have the matrix validated in order to assure that the diagram is correctly specified. Luckily, we already saw that it was the case earlier in this notebook:

```{r}
sfcr_sankey(tfm_pc, pc)
```

We can now move to simulation of different scenarios.


# Scenario 1: Increase in the rate of interest on bills

Let's start by increasing the rate of interest on bills by 100 points:

```{r}
shock1 <- sfcr_shock(
  variables = sfcr_set(
    r ~ 0.035
  ),
  start = 5,
  end = 60
)

pc2 <- sfcr_scenario(pc, scenario = shock1, periods = 60)
```


What happen in this scenario with the share of bills/money in household portfolios?

Since these ratios are not calculated when we simulate the scenario, we must calculate them now. We use the `dplyr` package from the `tidyverse` to manipulate the model:

```{r}
pc2 <- pc2 %>%
  mutate(BV = Bh / V,
         MV = Hh / V)

# If the read prefers to stick with base R functions, uncomment the following lines:
# pc2$BV <- pc2$B_h / pc2$V
# pc2$MV <- pc2$H_h / pc2$V
```

We reshape the model into the long format, and then we plot.

```{r}
pc2_long <- pc2 %>%
  pivot_longer(cols = -period)

pc2_long %>%
  filter(name %in% c("BV", "MV")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line() +
  facet_wrap(~ name, scales = 'free_y') +
  labs(title = "Wealth alocation")
```


```{r}
pc2_long %>%
  filter(name %in% c("YD", "C")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line(aes(linetype = name)) +
  labs(title = "Evolution of disposable income and household consumption",
       subtitle = "Following an increase of 100 points in the rate of interest")
```

## PCEX1: Random shocks to expected disposable income

To add a random shock to a variable with the `sfcr` package, one must create an auxiliary exogenous variable that will hold the exogenous shock, and add this exogenous variable to the endogenous variable of interest.

Let's see how to add a shock to expected disposable income, following equation 4.16 of @godley2007monetary.

First, we add the random shock, that we will call `Ra`, to the list of exogenous variables from the model `pc`. 

To make a normal random shock in R, we use the `rnorm()` function from the `stats` package that is built together with base R. This function accepts three arguments:

1. Number of observations;
2. Mean (default is 0);
3. Standard deviation (default is 1).

The user must set the same number of observations as the number of periods supplied to `sfcr_baseline()` or `sfcr_scenario()`. The other two arguments are optional.

```{r}
pc_ext2 <- c(pc_ext, Ra ~ rnorm(60, sd = 0.05))
```

Next, we rewrite the equations, adding the expectations to the model:

```{r}
pc_eqs2 <- pc_eqs

pc_eqs2[[5]] <- C ~ alpha1 * YDE + alpha2 * V[-1]
pc_eqs2[[8]] <- Bh ~ Bd
pc_eqs2[[12]] <- Hd ~ VE - Bd
pc_eqs2[[13]] <- VE ~ V[-1] + (YDE - C)
pc_eqs2[[14]] <- Bd ~ VE * (lambda0 + lambda1 * r - lambda2 * ( YDE/VE ))
pc_eqs2[[15]] <- YDE ~ YD * (1 + Ra)
```

We use the same parameters as before:

```{r}
pcex <- sfcr_baseline(pc_eqs2, pc_ext2, 60, hidden = c("Hh" = "Hs"))
```

The impact of adding these random noise to the expected disposable income can be seen by visualizing the money held balances vs the demand for money, and the changes in these variables:

```{r}
pcex <- pcex %>%
  mutate(delta_Hh = Hh - dplyr::lag(Hh),
         delta_Hd = Hd - dplyr::lag(Hd))

pcex_long <- pcex %>%
  pivot_longer(cols = -period)
```


```{r}
pcex_long %>%
  filter(name %in% c("Hh", "Hd")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line(aes(linetype = name)) +
  labs(title = "Evolution of stocks")
```

```{r}
pcex_long %>%
  filter(name %in% c("delta_Hh", "delta_Hd")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line(aes(linetype = name)) +
  labs(title = "Variation")
```

## PCEX2: Adaptive expectations

What would happen if, instead of random shocks to the expected disposable income, the households in our model followed a process of adaptive expectations?

```{r}
pc_eqs3 <- pc_eqs2
pc_eqs3[[15]] <- YDE ~ YD[-1]

pcex2 <- sfcr_baseline(pc_eqs3, pc_ext2, 60)
#
```

### Paradox of thrift?

One aspect of the model PCEX2 is that the paradox of thrift does not hold: an increase in the propensity to consume out of disposable income reduce the national income (Y) in the long run.

We can see this by adding a permanent increase to `alpha1` to model PCEX2:

```{r}
shock2 <- sfcr_shock(
  variables = sfcr_set(
    alpha1 ~ 0.7
  ),
  start = 5,
  end = 60
)

pc3 <- sfcr_scenario(pcex2, scenario = shock2, periods = 60)

pc3_long <- pc3 %>%
  #mutate(LV = dplyr::lag(V, n = 2L)) %>%
  pivot_longer(cols = -period)

```

```{r}
pc3_long %>%
  filter(name == "Y") %>%
  ggplot(aes(x = period, y = value)) +
  geom_line() +
  labs(title = "Evolution of National income")
```

```{r}
pc3_long %>%
  filter(name %in% c("C", "YDE", "V")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line(aes(linetype = name)) +
  labs(caption = "In the book it is written 'lagged' wealth but it is wealth actually.")
```

### The puzzling impact of interet rates reconsidered

What would happen with model PCEX2 if the propensity to consume (to save) becomes dependent on the rate of interest on bills?

```{r}
pc_eqs4 <- pc_eqs3
pc_eqs4[[16]] <- alpha1 ~ alpha10 - iota * r[-1]

pc_ext3 <- pc_ext

pc_ext3[[3]] <- alpha10 ~ 0.7
pc_ext3[[9]] <- iota ~ 4


pcex3 <- sfcr_baseline(pc_eqs4, pc_ext3, 60)
```


We add `shock1` to this model:

```{r}
pcex3_1 <- sfcr_scenario(pcex3, scenario = shock1, 60)

pcex3_1_long <- pcex3_1 %>%
  mutate(GExp = G + (dplyr::lag(r) * dplyr::lag(Bs)) - (dplyr::lag(r) * dplyr::lag(Bcb)) ) %>%
  pivot_longer(cols = -period)
```

```{r}
pcex3_1_long %>%
  filter(name %in% c("Y", "V", "YD", "C")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line(aes(linetype = name))
```


```{r}
pcex3_1_long %>%
  filter(name %in% c("GExp", "TX")) %>%
  ggplot(aes(x = period, y = value)) +
  geom_line(aes(linetype = name))
```


## Appendix: visualizing the sensitivity to parameters' values

This session requires a slightly more involved programming skills, and can be safely skipped by beginners.

When I was learning these models, I always wondered about the role of the parameters. For instance, in model PCEX2, one does only need a couple of trials to see that changing the parameters `alpha10` and `iota` will change the position of the lines. Let's explore this issue.

The `sfcr` package provides a set of functions that makes this type of analysis very easy. 

The most important function for this type of analysis is the `sfcr_multis()` function. This function generates multiple models from an expanded object and a fixed object. Currently, it accepts three possible combinations:

1. A `expanded` external set of parameters and exogenous values with a `fixed` set of model equations to generate multiple baseline models.

2. A `expanded` set of shocks that are applied to a `fixed` baseline model, generating multiple scenarios.

3. A `expanded` set of baseline models that are are applied to a `fixed` shock, generating multiple scenario models from multiple baseline models that receive the same shock.

In this notebook, we are going to explore the impact of different values for the `alpha10` by generating many baseline models and applying the same shock to all of these models.

We start by creating a vector with 20 possible values spanning from 0.61 to 0.8 withe the built-in `seq` function. We use this vector to expand the external set with the `sfcr_expand()` function.

```{r}
values = seq(0.61, 0.8, 0.01)
alpha10Exp <- sfcr_expand(x = pc_ext3, variable = alpha10, values = values)
#
```

We then use the `sfcr_multis()` function to generate multiple baseline models:

```{r}
pc_multi_bl <- sfcr_multis(alpha10Exp, pc_eqs4, periods = 50)
```

Finally, apply the `shock1` shock to all the models in `pc_multi_bl`:

```{r}
shock1 <- sfcr_shock(
  variables = sfcr_set(
    r ~ 0.035
  ),
  start = 5,
  end = 50
)

pc_multi_scn <- sfcr_multis(pc_multi_bl, shock1, periods = 50)
```

```{r, fig.width=9, fig.height=7}
# We also need to expand the color palettes
colourCount <- 20 # number of levels
getPalette <- grDevices::colorRampPalette(RColorBrewer::brewer.pal(6, "GnBu")[c(3:6)])


bind_rows(pc_multi_scn) %>%
  mutate(simulation = as_factor(simulation)) %>%
  pivot_longer(cols = -c(period, simulation)) %>%
  filter(name %in% c("Y", "YD", "C")) %>%
  ggplot(aes(x = period, y = value, color = simulation)) +
  geom_line() +
  theme(legend.position = "bottom") +
  scale_color_manual("alpha10",
    values = getPalette(20),
    labels = as.character(values)) +
  facet_wrap(~ name, scales = 'free_y')
```

We can see that the changing the `alpha10` parameter affects the level of the variables but do not change its underlying dynamics.

# References


